---
title: "Symptoms - scheduler experiment"
author: "Carlos Granell, SyMptOMS team, [GEOTEC research group](http://geotec.uji.es/) "
date: "17/4/2020 (updated `r format(Sys.time(), '%d %B, %Y')`)"
output: 
  html_document:
    df_print: paged
    toc: yes
---

```{r setup, include=FALSE, eval=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

This document analyses the self-developed sheduler componnet in Android-based mobile devices as part of the [SyMptOMS project](http://geotec.uji.es/projects/SyMptOMS/). 


## Reproduce paper

To create the PDF of the computational notebook you can run the following commands in a new R session.
If you have problems rendering the PDF you can execute each chunk independently in [RStudio](https://rstudio.com/products/rstudio/).

```{r render_with_rmarkdown, eval=FALSE}
require("knitr")
require("rmarkdown")
rmarkdown::render("symptoms-scheduler-experiment.Rmd", output_format = "pdf_document")
```

This document does not install the required R packages by default.
You can run the script `install.R` to install all required dependencies on a new R installation, or use `install.packages(..)` to install missing R packages.

```{r install_r, eval=FALSE}
source("R/install.R")
```

The plots and tables use the packages [`ggplot2`](http://ggplot2.tidyverse.org/), [`knitr::kable()`](https://yihui.name/knitr/) and [`kableExtra`](https://cran.r-project.org/package=kableExtra).

Required libraries and runtime environment description are as follows.

```{r load_libraries, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(kableExtra)
library(here)
library(googledrive)
library(lubridate)
library(stringr)
```


```{r set_seed}
# just in case
set.seed(nchar("Symptoms - scheduler experiment"))
```

```{r gsheetdata_files}

files_per_experiment <- 5

files <- tibble(
  experiment = c(rep("#1", files_per_experiment), rep("#2", files_per_experiment)),
  gsheets_name = c("AD_BQ.csv", "AD_NV.csv", "AD_A1.csv", "BA_H9.csv", "BA_MO.csv",
                   "BA_BQ.csv", "BA_NV.csv", "BA_A1.csv", "AD_H9.csv", "AD_MO.csv")
  #TODO: issue readinf private files from google drive: files cannot be downloaded    
  # gsheets_id = c() 
)
```



```{r drive_connect, eval=FALSE}
gfolder_url <- "https://drive.google.com/open?id=11oqV_vZqRDkbMdQ8m2KmAeL3-mOPKClh"
gdata_path <- drive_get(as_id(gfolder_url))
```

```{r gsheetdata_exp1_download, eval=FALSE}
# Download files corresponding to 'Experiment #1'
files_exp <- files %>%
  filter(experiment =="#1") %>%
  select(gsheets_name)

for (f in 1:files_per_experiment) {
  gfile_name <-files_exp$gsheets_name[f]
  gdata_file <- drive_ls(path = gdata_path$name, pattern = gfile_name)

  data_path <- here::here("data-raw", gdata_file$name) # local file
  drive_download(file = as_id(gdata_file$id), path = data_path, overwrite = TRUE, verbose = TRUE)
}

```

```{r gsheetdata_exp2_download, eval=FALSE}
# Download files corresponding to 'Experiment #2'
files_exp <- files %>%
  filter(experiment =="#2") %>%
  select(gsheets_name)

for (f in 1:files_per_experiment) {
  gfile_name <-files_exp$gsheets_name[f]
  gdata_file <- drive_ls(path = gdata_path$name, pattern = gfile_name)

  data_path <- here::here("data-raw", gdata_file$name) # local file
  drive_download(file = as_id(gdata_file$id), path = data_path, overwrite = TRUE, verbose = TRUE)
}

```

```{r drive_disconnect, eval=FALSE}
drive_deauth()
```


## Data preparation

```{r merge_datafiles}
source(here::here("R", "02_merge_datafiles.R"))
```

## Scheduler assessment

```{r load_assessdata, warning=FALSE}
assessment_file <- here::here("data", "data.rds")
data <- readRDS(assessment_file)
```

```{r summary_table}

total_observations <- nrow(data) 

summaries <- 
  data %>%
  group_by(exp_id, device_desc) %>%
  summarise(`# observations` = n(),
            `start datetime` = min(plan_date),
            `end_datetime` = max(plan_date),
            `mean delay` = first(mean),
            `sd` = first(sd),
            `mean - 2sd` = first(lo),
            `mean + 2sd` = first(hi))

kable(summaries,
      format = "html",
      booktabs = TRUE,
      caption = paste("Information of data collected per device. Total number of observations =", total_observations)) %>%
  kable_styling(full_width =TRUE)

 

```



