---
title: "Symptoms - 'NativeScript Task Dispatcher' experiment (I)"
author: "Carlos Granell, [SyMptOMS team](http://geotec.uji.es/projects/SyMptOMS/), [GEOTEC research group](http://geotec.uji.es/) "
date: "17/4/2020 (updated `r format(Sys.time(), '%d %B, %Y')`)"
output: 
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 4
---

```{r setup, include=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo = FALSE)
```

This document analyses the self-developed scheduler component in Android-based mobile devices as part of the [SyMptOMS project](http://geotec.uji.es/projects/SyMptOMS/). This sofware component tested here is called [NativeScript Task Dispatcher](https://www.npmjs.com/package/nativescript-task-dispatcher), that's a [NativeScript](https://nativescript.org/) plugin aimed to ease mobile app's tasks definition and the execution of task workflows in background, regardless of whether the app is visible to the user or not. [NativeScript Task Dispatcher](https://www.npmjs.com/package/nativescript-task-dispatcher) is publicy available as open source software. 


## Reproduce notebook

To create the PDF of the computational notebook you can run the following commands in a new R session.
If you have problems rendering the PDF you can execute each chunk independently in [RStudio](https://rstudio.com/products/rstudio/).

```{r render_with_rmarkdown, eval=FALSE}
require("knitr")
require("rmarkdown")
rmarkdown::render("symptoms-scheduler-experiment-1.Rmd", output_format = "pdf_document")
```

This document does not install the required R packages by default.
You can run the script `install.R` to install all required dependencies on a new R installation, or use `install.packages(..)` to install missing R packages.

```{r install_r, eval=FALSE}
source("R/install.R")
```

The plots and tables use the packages [`ggplot2`](http://ggplot2.tidyverse.org/), [`knitr::kable()`](https://yihui.name/knitr/), [`kableExtra`](https://cran.r-project.org/package=kableExtra) and [`timetk`](https://business-science.github.io/timetk/index.html).

Required libraries and runtime environment description are as follows.

```{r load_libraries, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(kableExtra)
library(here)
# library(googledrive)
library(lubridate)
library(stringr)
library(scales)
# library(plotly)
library(grid)
library(gridBase)
library(gridExtra)
# library(egg)
library(timetk)
```


```{r set_seed}
# just in case
set.seed(nchar("Symptoms - 'NativeScript Task Dispatcher' experiment"))
```

## Data download

Log files are stored in shared gdrvie folder. This notebook downloads them locally in the folder `data-raw`.

```{r gsheetdata_files}

files <- tibble(
  experiment = c(rep("#1", 5), rep("#2", 5)),
  gsheets_name = c("AD_BQ.csv", "AD_NV.csv", "AD_A1.csv", "BA_H9.csv", "BA_MO.csv",
                   "BA_BQ.csv", "BA_NV.csv", "BA_A1.csv", "AD_H9.csv", "AD_MO.csv"),
  gsheets_link = c("https://drive.google.com/open?id=1fP9r0S8ORa689yHRtfCtYQothLuGcCw3",
                   "https://drive.google.com/open?id=1l-vnaT-Smy0SegArIz5A108-IQ2_jmVV",
                   "https://drive.google.com/open?id=1gzk1ezN5t5yBv2RnbznUqvalDx97eu3o",
                   "https://drive.google.com/open?id=1BLXO4Rvz6ppWFJe5oN9Y0hKh__ypofxk",
                   "https://drive.google.com/open?id=1RizGjKR8QLOdJna7qGPP7dYX9WWAysO4",
                   "https://drive.google.com/open?id=1rfaooLZ0Up0gNx1SCA92TV1JFxcLU0sk",
                   "https://drive.google.com/open?id=1_KIXX5FGXv7MO37QbUDnd_KGqiGihiMI",
                   "https://drive.google.com/open?id=1LezrtyB9i4K7cQsmI78E0v_KaOyhZcXI",
                   "https://drive.google.com/open?id=1ISjsfSLZt6n9miryQtDr-tsZV9bfbCDT",
                   "https://drive.google.com/open?id=1bx2RVNz05qVKGr-32GI-ZIOKQWlqKBPz")
)
```

```{r gsheetdata_files_table}
kable(files,
      format = "html",
      booktabs = TRUE,
      caption = "Table. List of raw data files.") %>%
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "bordered", "condensed"))

```


```{r gsheetdata_download, eval=FALSE}
for (f in 1:nrow(files)) {
  gfile_name <-files$gsheets_name[f]
  gfile_id <- googledrive::as_id(files$gsheets_link[f])
  
  drive_download(file = gfile_id, 
                 path = here::here("data-raw", gfile_name), 
                 overwrite = TRUE, verbose = FALSE)
}

```

```{r drive_disconnect, eval=FALSE}
drive_deauth()
```


## Data preparation

Raw data files are cleaned and merged together into a single file stored in the folder `data`. New fields such as dates breakdown are computed too. 

```{r merge_datafiles_defs}

schedulers <- c("AD" = "Advanced", 
                "BA" = "Basic")

devices <- c("BQ" = "BQ Aquaris V",
             "NV" = "Nvidia Shield Tablet",
             "A1" = "Xiaomi Mi A1",
             "H9" = "Honor 9",
             "MO" = "Motorola Moto G")

baseline_delay <- 60
```


```{r merge_datafiles, eval=FALSE}

data_path <- here::here("data-raw")
file_names <- list.files(path  = data_path)
file_paths <- list.files(path = data_path, full.names = TRUE)

data_merged <- data.frame()

get_experiment_id <- function(filename) {
  fullname <- paste0(filename, ".csv")
  files %>%
    filter(gsheets_name == fullname) %>%
    select(experiment) %>%
    pull
}

for (f in 1:length(file_names)) {
  filename <- stringr::str_sub(file_names[f], 1, 5)
  exp_id <- get_experiment_id(filename)
  
  parts <- stringr::str_split(filename, "_", simplify = TRUE)
  
  scheduler_id = parts[1]
  scheduler_name <- schedulers[[scheduler_id]]
  device_id <- parts[2]
  device_name <- devices[[device_id]]
  device_desc <- paste0(scheduler_name, " - ", device_name)
  
  data_temp <- read_csv(file_paths[f], col_names = TRUE, 
                        cols(
                          battery = col_double(),
                          exec_timestamp = col_double(),
                          planning_timestamp = col_double(),
                          task = col_character()
                        ))


  data_temp <- 
    data_temp %>%
    mutate(exp_id = exp_id,
           scheduler = scheduler_name,
           device_id = device_id,
           device_name = device_name,
           device_desc = device_desc) 
  
  data_merged <- rbind(data_merged, data_temp)
}       


data_merged <- 
  data_merged %>%
  mutate(exec_date = as_datetime(exec_timestamp/1000, tz="Europe/Madrid"),
         plan_date = as_datetime(planning_timestamp/1000, tz="Europe/Madrid"),
         plan_minute = lubridate::minute(plan_date),
         plan_hour = lubridate::hour(plan_date),
         plan_day = lubridate::day(plan_date),
         plan_month = lubridate::month(plan_date),
         exec_minute = lubridate::minute(exec_date),
         exec_hour = lubridate::hour(exec_date),
         exec_day = lubridate::day(exec_date),
         exec_month = lubridate::month(exec_date),
         time_period = ifelse(between(plan_hour,0, 7), "nighttime", "daytime"))


# Delay units: seconds
data_merged <-
  data_merged %>%
  group_by(device_id) %>%
  arrange(plan_date) %>%
  mutate(step = row_number(),
         delay = (as.duration(interval(plan_date, exec_date)) - baseline_delay) / dseconds(1))

data_path <- here::here("data", "data.csv")
write_csv(data_complete, data_path)
data_path <- here::here("data", "data.rds")
saveRDS(data_complete, data_path)

#TODO: shiny app is deprecated
# data_path <- here::here("webapp","data", "data.csv")
# write_csv(data_complete, data_path)
# data_path <- here::here("webapp","data", "data.rds")
# saveRDS(data_complete, data_path)

```

__Key Variables__:

- quantitative (interval)

  - `plan_date` (_datetime_): task planning time.
  - `exec_date` (_datetime_): task execution time.  

- quantitative (ratio)

  - `delay` (_numeric_): execution delay in seconds ((`exec_date` - `plan_date`) - `r baseline_delay`).

- categorital (ordinal)

  - `battery`(_numeric_): battery level

- categorical (nominal)

  - `device_id`/`device_name`: device identifier/name
  - `exp_id`: experiment identifier
  - `scheduler`: scheduler type
  

## Exploratory analysis


### Summary statistics

```{r load_assessdata, warning=FALSE}
assessment_file <- here::here("data", "data.rds")
data <- readRDS(assessment_file)

n_observations <- nrow(data) 

```


```{r stats_summary_table, warning=FALSE, message=FALSE}

# https://towardsdatascience.com/understanding-boxplots-5e2df7bcbd51
# Understanding Boxplot. What is a outlier?
# Min., 1st Qu., Median, Mean, 3rd Qu., Max.

col_names <- c("scheduler", "device_name", "min", "quartile1", "median", "quartile3", "max", "mean", "bottom_whisker", "top_whisker")
summaries <- data.frame(matrix(ncol=length(col_names), nrow = 0))
colnames(summaries) <- col_names

for (s in 1:length(schedulers)) {
  for (d in 1:length(devices)) {
    
    data %>%
      filter(scheduler == schedulers[[s]]) %>%
      filter(device_name == devices[[d]]) %>%
      select(device_id, device_name, delay) -> one
    
    summary_row <- data.frame(matrix(ncol=length(col_names), nrow = 0))
    colnames(summary_row) <- col_names
    
    mean <- mean(one$delay)
    summary_boxplot <- boxplot(one$delay, plot = FALSE) 
    bottom_whisker <- summary_boxplot$stats[1,1]
    top_whisker <-  summary_boxplot$stats[5,1]
    
    summary_row[1,] <- c(schedulers[[s]], devices[[d]], fivenum(one$delay), mean, bottom_whisker, top_whisker)

    summaries <- rbind(summaries, summary_row)
  }
}

summaries <- 
  summaries %>%
  mutate_at(vars(c("scheduler", "device_name")), ~as.character(.)) %>% 
  mutate_at(vars(-c("scheduler", "device_name")), ~as.numeric(.)) %>%
  mutate_if(is.numeric, ~round(., 3))

# glimpse(summaries)

data %>%
  group_by(exp_id, scheduler, device_name) %>%
  summarise(n = n(),
            start_datetime = min(plan_date),
            end_datetime = max(plan_date)) %>%
  inner_join(summaries, by=c("scheduler", "device_name")) %>%
  arrange(exp_id, scheduler, device_name) -> stats_summaries

kable(stats_summaries %>%
        select(`# exp` = exp_id,
               `Type scheduler` = scheduler,
               `Device` = device_name,
               `# records` = n,
               `Start` = start_datetime,
               `End` = end_datetime,
               `Min` = min,
               `Bottom whisker`= bottom_whisker,
               `1st Qu.` = quartile1,
               `Median` = median,
               `Mean` = mean,
               `3rd Qu.` = quartile3,
               `Top whisker`= top_whisker,
               `Max` = max),
      format = "html",
      booktabs = TRUE,
      caption = "Table. Descriptive statistics of data distribution and centrality per scheduler/device.") %>%
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "bordered", "condensed")) %>%
  collapse_rows(columns = 1:2, valign = "top") %>%
  # row_spec(6, color = "white", background = "orange") %>%
  row_spec(9:10, color = "white", background = "#D7261E")

```



```{r outliers}

# devices_OK <- c("BQ" = "BQ Aquaris V",
#                 "H9" = "Honor 9",
#                 "MO" = "Motorola Moto G")

data %>% 
  left_join(summaries, by=c("scheduler", "device_name")) %>%
  mutate(outlier_pos = ifelse(delay > top_whisker, 1, 0)) %>%
  mutate(outlier_neg = ifelse(delay < bottom_whisker, 1, 0)) %>%
  mutate(outlier = ifelse((outlier_pos == 1 | outlier_neg == 1), 1, 0)) %>%
  mutate(whisker = ifelse((delay > bottom_whisker & delay < quartile1) | (delay > quartile3 & delay < top_whisker), 1, 0)) %>%
  select(-min,-bottom_whisker, -quartile1, -mean, -median, -quartile3, -top_whisker, -max) %>%
  arrange(device_id, plan_date) -> data


n_outliers <- nrow(filter(data, outlier == 1))
n_whisker <- nrow(filter(data, whisker == 1))
n_irq <- nrow(filter(data, outlier == 0 & whisker == 0))

```

**Observations on the above table **:

- `r scales::comma(n_observations, big.mark = ".")` records in total. 

- `r round((n_irq / n_observations)*100, 1)`% (`r scales::comma(n_irq, big.mark = ".")`/`r scales::comma(n_observations, big.mark = ".")`) is the interquartile range or [IQR](https://en.wikipedia.org/wiki/Interquartile_range), i.e. the middle 50% of the data.

- `r round((n_whisker / n_observations)*100, 1)`% (`r scales::comma(n_whisker, big.mark = ".")`/`r scales::comma(n_observations, big.mark = ".")`) represent the _reasonable_ extremes of the data, i.e. these whisker values are the minimum and maximum values that do not exceed 1.5 × IQR.

- `r round((n_outliers / n_observations)*100, 1)`% (`r scales::comma(n_outliers, big.mark = ".")`/`r scales::comma(n_observations, big.mark = ".")`) are outliers, i.e. values beyond the top and bottom whiskers.

- `Nvidia Shield Tablet` has stopped triggering alarms on  **2020-04-17 22:27:06** when running the basic scheduler. Since there's no _watchdog_ in the basic version, this device is not going to trigger alarms anymore. Died!

- `Xiaomi Mi A1` has stopped triggering alarms on **2020-04-23 00:52:00** when running the basic scheduler. An explanation may be that Huawei/Honor smartphones on Android 9+ have a Launch Records where the OS records how much times an app has been launched by itself, so the OS has the possibility to detect this kind of apps. In these smartphones, Huawei also introduced a new task killer (Power Genius) which kills everything that is not whitelisted by Huawei, and does not give users any configuration options. Therefore, the Power Genius uses the information from the Launch Records to kill the scheduled alarms




### Time series visualisation (Distribution of `delay`)

```{r ts_defs}
# Setup for the plotly charts (# FALSE returns ggplots)
interactive <- TRUE 
```

#### BQ Aquaris V

```{r ts_delay_BQ_AD, fig.height=8, fig.width=12, warning=FALSE, message=FALSE}

data %>%
  filter(scheduler==schedulers[[1]], device_name==devices[[1]]) %>%
  select(scheduler, device_name, plan_date, delay) %>%
  plot_time_series(plan_date, delay,
                   # .facet_ncol = 1, .facet_scales = "free",
                   .color_var = week(plan_date),
                   .interactive = interactive,
                   .title = paste("[Delay]", schedulers[[1]], "-", devices[[1]]),
                   .y_lab = "Delay [seconds]",
                   .color_lab = "Week", 
                   .plotly_slider = TRUE)

```



```{r ts_delay_BQ_BA, fig.height=8, fig.width=12, warning=FALSE, message=FALSE}

data %>%
  filter(scheduler==schedulers[[2]], device_name==devices[[1]]) %>%
  select(scheduler, device_name, plan_date, delay) %>%
  plot_time_series(plan_date, delay,
                   # .facet_ncol = 1, .facet_scales = "free",
                   .color_var = week(plan_date),
                   .interactive = interactive,
                   .title = paste("[Delay]", schedulers[[2]], "-", devices[[1]]),
                   .y_lab = "Delay [seconds]",
                   .color_lab = "Week", 
                   .plotly_slider = TRUE)


```


```{r ts_delay_BQ_seasonality, fig.height=16, fig.width=12}

data %>%
  filter(device_name==devices[[1]]) %>%
  group_by(scheduler) %>%
  select(scheduler, device_name, plan_date, delay) %>%
    plot_seasonal_diagnostics(plan_date, delay, 
                              .title = paste("[Seasonal Diagnostics]", devices[[1]]),
                              .interactive = interactive)

```


#### Nvidia Shield Tablet


```{r ts_delay_NV_AD, fig.height=8, fig.width=12, warning=FALSE, message=FALSE}

data %>%
  filter(scheduler==schedulers[[1]], device_name==devices[[2]]) %>%
  select(scheduler, device_name, plan_date, delay) %>%
  plot_time_series(plan_date, delay,
                   # .facet_ncol = 1, .facet_scales = "free",
                   .color_var = week(plan_date),
                   .interactive = interactive,
                   .title = paste("[Delay]", schedulers[[1]], "-", devices[[2]]),
                   .y_lab = "Delay [seconds]",
                   .color_lab = "Week", 
                   .plotly_slider = TRUE)

```



```{r ts_delay_NV_BA, fig.height=8, fig.width=12, warning=FALSE, message=FALSE}

data %>%
  filter(scheduler==schedulers[[2]], device_name==devices[[2]]) %>%
  select(scheduler, device_name, plan_date, delay) %>%
  plot_time_series(plan_date, delay,
                   # .facet_ncol = 1, .facet_scales = "free",
                   .color_var = week(plan_date),
                   .interactive = interactive,
                   .title = paste("[Delay]", schedulers[[2]], "-", devices[[2]]),
                   .y_lab = "Delay [seconds]",
                   .color_lab = "Week", 
                   .plotly_slider = TRUE)


```


```{r ts_delay_NV_seasonality, fig.height=16, fig.width=12}

data %>%
  filter(device_name==devices[[2]]) %>%
  group_by(scheduler) %>%
  select(scheduler, device_name, plan_date, delay) %>%
    plot_seasonal_diagnostics(plan_date, delay, 
                              .title = paste("[Seasonal Diagnostics]", devices[[2]]),
                              .interactive = interactive)

```




#### Xiaomi Mi A1




```{r ts_delay_A1_AD, fig.height=8, fig.width=12, warning=FALSE, message=FALSE}

data %>%
  filter(scheduler==schedulers[[1]], device_name==devices[[3]]) %>%
  select(scheduler, device_name, plan_date, delay) %>%
  plot_time_series(plan_date, delay,
                   # .facet_ncol = 1, .facet_scales = "free",
                   .color_var = week(plan_date),
                   .interactive = interactive,
                   .title = paste("[Delay]", schedulers[[1]], "-", devices[[3]]),
                   .y_lab = "Delay [seconds]",
                   .color_lab = "Week", 
                   .plotly_slider = TRUE)

```



```{r ts_delay_A1_BA, fig.height=8, fig.width=12, warning=FALSE, message=FALSE}

data %>%
  filter(scheduler==schedulers[[2]], device_name==devices[[3]]) %>%
  select(scheduler, device_name, plan_date, delay) %>%
  plot_time_series(plan_date, delay,
                   # .facet_ncol = 1, .facet_scales = "free",
                   .color_var = week(plan_date),
                   .interactive = interactive,
                   .title = paste("[Delay]", schedulers[[2]], "-", devices[[3]]),
                   .y_lab = "Delay [seconds]",
                   .color_lab = "Week", 
                   .plotly_slider = TRUE)


```


```{r ts_delay_A1_seasonality, fig.height=16, fig.width=12}

data %>%
  filter(device_name==devices[[3]]) %>%
  group_by(scheduler) %>%
  select(scheduler, device_name, plan_date, delay) %>%
    plot_seasonal_diagnostics(plan_date, delay, 
                              .title = paste("[Seasonal Diagnostics]", devices[[3]]),
                              .interactive = interactive)

```




#### Honor 9




```{r ts_delay_H9_AD, fig.height=8, fig.width=12, warning=FALSE, message=FALSE}

data %>%
  filter(scheduler==schedulers[[1]], device_name==devices[[4]]) %>%
  select(scheduler, device_name, plan_date, delay) %>%
  plot_time_series(plan_date, delay,
                   # .facet_ncol = 1, .facet_scales = "free",
                   .color_var = week(plan_date),
                   .interactive = interactive,
                   .title = paste("[Delay]", schedulers[[1]], "-", devices[[4]]),
                   .y_lab = "Delay [seconds]",
                   .color_lab = "Week", 
                   .plotly_slider = TRUE)

```



```{r ts_delay_H9_BA, fig.height=8, fig.width=12, warning=FALSE, message=FALSE}

data %>%
  filter(scheduler==schedulers[[2]], device_name==devices[[4]]) %>%
  select(scheduler, device_name, plan_date, delay) %>%
  plot_time_series(plan_date, delay,
                   # .facet_ncol = 1, .facet_scales = "free",
                   .color_var = week(plan_date),
                   .interactive = interactive,
                   .title = paste("[Delay]", schedulers[[2]], "-", devices[[4]]),
                   .y_lab = "Delay [seconds]",
                   .color_lab = "Week", 
                   .plotly_slider = TRUE)


```


```{r ts_delay_H9_seasonality, fig.height=16, fig.width=12}

data %>%
  filter(device_name==devices[[4]]) %>%
  group_by(scheduler) %>%
  select(scheduler, device_name, plan_date, delay) %>%
    plot_seasonal_diagnostics(plan_date, delay, 
                              .title = paste("[Seasonal Diagnostics]", devices[[4]]),
                              .interactive = interactive)

```


#### Motorola Moto G






```{r ts_delay_MO_AD, fig.height=8, fig.width=12, warning=FALSE, message=FALSE}

data %>%
  filter(scheduler==schedulers[[1]], device_name==devices[[5]]) %>%
  select(scheduler, device_name, plan_date, delay) %>%
  plot_time_series(plan_date, delay,
                   # .facet_ncol = 1, .facet_scales = "free",
                   .color_var = week(plan_date),
                   .interactive = interactive,
                   .title = paste("[Delay]", schedulers[[1]], "-", devices[[5]]),
                   .y_lab = "Delay [seconds]",
                   .color_lab = "Week", 
                   .plotly_slider = TRUE)

```



```{r ts_delay_MO_BA, fig.height=8, fig.width=12, warning=FALSE, message=FALSE}

data %>%
  filter(scheduler==schedulers[[2]], device_name==devices[[5]]) %>%
  select(scheduler, device_name, plan_date, delay) %>%
  plot_time_series(plan_date, delay,
                   # .facet_ncol = 1, .facet_scales = "free",
                   .color_var = week(plan_date),
                   .interactive = interactive,
                   .title = paste("[Delay]", schedulers[[2]], "-", devices[[5]]),
                   .y_lab = "Delay [seconds]",
                   .color_lab = "Week", 
                   .plotly_slider = TRUE)


```


```{r ts_delay_MO_seasonality, fig.height=16, fig.width=12}

data %>%
  filter(device_name==devices[[5]]) %>%
  group_by(scheduler) %>%
  select(scheduler, device_name, plan_date, delay) %>%
    plot_seasonal_diagnostics(plan_date, delay, 
                              .title = paste("[Seasonal Diagnostics]", devices[[5]]),
                              .interactive = interactive)

```





```{r ts_basic_scheduler, fig.height=16, fig.width=12, eval=FALSE}

# data %>%
#   filter(scheduler==schedulers[[2]]) %>%
#   group_by(device_name) %>%
#   select(device_name, plan_date, delay) %>%
#   plot_time_series(plan_date, delay,
#                    .facet_ncol = 1, .facet_scales = "free",
#                    .interactive = interactive,
#                    .title = schedulers[[2]],
#                     .y_lab = "Delay [seconds]",
#                    .plotly_slider = TRUE)
```



