---
title: "Symptoms - scheduler experiment timestamps (Part II)"
author: "Carlos Granell, [SyMptOMS team](http://geotec.uji.es/projects/SyMptOMS/), [GEOTEC research group](http://geotec.uji.es/) "
date: "21/4/2020 (updated `r format(Sys.time(), '%d %B, %Y')`)"
output: 
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 4
---

```{r setup, include=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo = FALSE)
```

This document analyses the self-developed scheduler component in Android-based mobile devices as part of the [SyMptOMS project](http://geotec.uji.es/projects/SyMptOMS/). 


## Reproduce notebook

To create the PDF of the computational notebook you can run the following commands in a new R session.
If you have problems rendering the PDF you can execute each chunk independently in [RStudio](https://rstudio.com/products/rstudio/).

```{r render_with_rmarkdown, eval=FALSE}
require("knitr")
require("rmarkdown")
rmarkdown::render("symptoms-scheduler-experiment-timestamps.Rmd", output_format = "pdf_document")
```

This document does not install the required R packages by default.
You can run the script `install.R` to install all required dependencies on a new R installation, or use `install.packages(..)` to install missing R packages.

```{r install_r, eval=FALSE}
source("R/install.R")
```

The plots and tables use the packages [`ggplot2`](http://ggplot2.tidyverse.org/), [`knitr::kable()`](https://yihui.name/knitr/) and [`kableExtra`](https://cran.r-project.org/package=kableExtra).

Required libraries and runtime environment description are as follows.

```{r load_libraries, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(kableExtra)
library(here)
# library(googledrive)
library(lubridate)
library(stringr)
library(scales)
# library(plotly)
library(grid)
library(gridBase)
library(gridExtra)
library(egg)
```


```{r set_seed}
# just in case
set.seed(nchar("Symptoms - scheduler experiment timestamps"))
```

## Data download

Log files are stored in shared gdrvie folder. This notebook downloads them locally in the folder`data-raw`.

```{r gsheetdata_files}

files <- tibble(
  experiment = c(rep("#1", 5), rep("#2", 5)),
  gsheets_name = c("AD_BQ.csv", "AD_NV.csv", "AD_A1.csv", "BA_H9.csv", "BA_MO.csv",
                   "BA_BQ.csv", "BA_NV.csv", "BA_A1.csv", "AD_H9.csv", "AD_MO.csv"),
  gsheets_link = c("https://drive.google.com/open?id=1fP9r0S8ORa689yHRtfCtYQothLuGcCw3",
                   "https://drive.google.com/open?id=1l-vnaT-Smy0SegArIz5A108-IQ2_jmVV",
                   "https://drive.google.com/open?id=1gzk1ezN5t5yBv2RnbznUqvalDx97eu3o",
                   "https://drive.google.com/open?id=1BLXO4Rvz6ppWFJe5oN9Y0hKh__ypofxk",
                   "https://drive.google.com/open?id=1RizGjKR8QLOdJna7qGPP7dYX9WWAysO4",
                   "https://drive.google.com/open?id=1rfaooLZ0Up0gNx1SCA92TV1JFxcLU0sk",
                   "https://drive.google.com/open?id=1_KIXX5FGXv7MO37QbUDnd_KGqiGihiMI",
                   "https://drive.google.com/open?id=1LezrtyB9i4K7cQsmI78E0v_KaOyhZcXI",
                   "https://drive.google.com/open?id=1ISjsfSLZt6n9miryQtDr-tsZV9bfbCDT",
                   "https://drive.google.com/open?id=1bx2RVNz05qVKGr-32GI-ZIOKQWlqKBPz")
)
```

```{r gsheetdata_files_table}
kable(files,
      format = "html",
      booktabs = TRUE,
      caption = "Table. List of raw data files.") %>%
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "bordered", "condensed"))

```


```{r gsheetdata_download, eval=FALSE}
for (f in 1:nrow(files)) {
  gfile_name <-files$gsheets_name[f]
  gfile_id <- googledrive::as_id(files$gsheets_link[f])
  
  drive_download(file = gfile_id, 
                 path = here::here("data-raw", gfile_name), 
                 overwrite = TRUE, verbose = FALSE)
}

```

```{r drive_disconnect, eval=FALSE}
drive_deauth()
```


## Data preparation

Raw data files are cleaned and merged together into a single file stored in the folder `data`. New fields such as dates breakdown are computed too. 

```{r merge_datafiles_defs}

schedulers <- c("AD" = "Advanced", 
                "BA" = "Basic")

devices <- c("BQ" = "BQ Aquaris V",
             "NV" = "Nvidia Shield Tablet",
             "A1" = "Xiaomi Mi A1",
             "H9" = "Honor 9",
             "MO" = "Motorola Moto G")

baseline_delay <- 60
```


```{r merge_datafiles, eval=FALSE}

data_path <- here::here("data-raw")
file_names <- list.files(path  = data_path)
file_paths <- list.files(path = data_path, full.names = TRUE)

data_merged <- data.frame()

get_experiment_id <- function(filename) {
  fullname <- paste0(filename, ".csv")
  files %>%
    filter(gsheets_name == fullname) %>%
    select(experiment) %>%
    pull
}

for (f in 1:length(file_names)) {
  filename <- stringr::str_sub(file_names[f], 1, 5)
  exp_id <- get_experiment_id(filename)
  
  parts <- stringr::str_split(filename, "_", simplify = TRUE)
  
  scheduler_id = parts[1]
  scheduler_name <- schedulers[[scheduler_id]]
  device_id <- parts[2]
  device_name <- devices[[device_id]]
  device_desc <- paste0(scheduler_name, " - ", device_name)
  
  data_temp <- read_csv(file_paths[f], col_names = TRUE, 
                        cols(
                          battery = col_double(),
                          exec_timestamp = col_double(),
                          planning_timestamp = col_double(),
                          task = col_character()
                        ))


  data_temp <- 
    data_temp %>%
    mutate(exp_id = exp_id,
           scheduler = scheduler_name,
           device_id = device_id,
           device_name = device_name,
           device_desc = device_desc) 
  
  data_merged <- rbind(data_merged, data_temp)
}       


data_merged <- 
  data_merged %>%
  mutate(exec_date = as_datetime(exec_timestamp/1000, tz="Europe/Madrid"),
         plan_date = as_datetime(planning_timestamp/1000, tz="Europe/Madrid"),
         plan_minute = lubridate::minute(plan_date),
         plan_hour = lubridate::hour(plan_date),
         plan_day = lubridate::day(plan_date),
         plan_month = lubridate::month(plan_date),
         exec_minute = lubridate::minute(exec_date),
         exec_hour = lubridate::hour(exec_date),
         exec_day = lubridate::day(exec_date),
         exec_month = lubridate::month(exec_date),
         time_period = ifelse(between(plan_hour,0, 7), "nighttime", "daytime"))


# Delay units: seconds
data_merged <-
  data_merged %>%
  group_by(scheduler, device_id) %>%
  arrange(plan_date) %>%
  mutate(step = row_number(),
         delay = (as.duration(interval(plan_date, exec_date)) - baseline_delay) / dseconds(1))


data_path <- here::here("data", "data.csv")
write_csv(data_complete, data_path)
data_path <- here::here("data", "data.rds")
saveRDS(data_complete, data_path)

#TODO: shiny app is deprecated
# data_path <- here::here("webapp","data", "data.csv")
# write_csv(data_complete, data_path)
# data_path <- here::here("webapp","data", "data.rds")
# saveRDS(data_complete, data_path)

```

__Key Variables__:

- quantitative (interval)

  - `plan_date` (_datetime_): task planning time.
  - `exec_date` (_datetime_): task execution time.  
- quantitative (ratio)

  - `delay` (_numeric_): execution delay in seconds ((`exec_date` - `plan_date`) - `r baseline_delay`).
- categorital (ordinal)

  - `battery`(_numeric_): battery level
- categorical (nominal)

  - `device_id`/`device_name`: device identifier/name
  - `exp_id`: experiment identifier.
  - `scheduler`: scheduler type.
  

## Exploratory analysis


```{r load_assessdata, warning=FALSE}
assessment_file <- here::here("data", "data.rds")
data <- readRDS(assessment_file)

n_observations <- nrow(data) 

```


### Which one of the analysed schedulers is more accurate? (Distribution of `delay`)


```{r dual_scale_plots_function}
source(here::here("R", "dual_scale_plot.R"))
```

Basic assumption: Higher accuracy = lower `delay`. But what is delay? We compute the following timestamps diffs: 

- `planned timestamp(n-1) - planned timestamp(n)`, in order to detect wrong planned timestamp values 

- `execution timestamp(n-1) - execution timestamp(n)` , to detect execution timestamps betwwen consecutive observations

- execution timestamp(n) - planning timestamp(n+1) [the current one, which displays framework overhead]



#### Analsysis of consecutive planning timestamps 

We calculate `planned timestamp(n-1) - planned timestamp(n)` to detect wrong planning timestamp values. During data collection, chances are that some planned timestamps are not correctly registered, creating duplicated timestamps, suggested  by the presence of negative values of `delay` 

```{r plan_dates_diff}
diff_times <- 
  data %>%
  select(step, scheduler, device_id, device_name, delay, plan_date, exec_date, plan_day, exec_day)

diff_times <- 
  diff_times %>%
  group_by(scheduler, device_id) %>%
  arrange(plan_date) %>%
  mutate(plan_date_lagged = lag(plan_date),
         plan_date_diff = as.numeric(as.duration(interval(plan_date_lagged, plan_date)) / dseconds(1)),
         plan_date_diff = ifelse(is.na(plan_date_diff), 0, round(plan_date_diff, 3)),
         plan_date_diff_cum = cumsum(plan_date_diff))
         
```


```{r plan_dates_diff_summary, warning=FALSE}

# https://towardsdatascience.com/understanding-boxplots-5e2df7bcbd51
# Understanding Boxplot. What is a outlier?
# Min., 1st Qu., Median, Mean, 3rd Qu., Max.

col_names <- c("scheduler", "device_name", "min", "quartile1", "median", "quartile3", "max", "mean", "bottom_whisker", "top_whisker")
summaries_diff <- data.frame(matrix(ncol=length(col_names), nrow = 0))
colnames(summaries_diff) <- col_names

for (s in 1:length(schedulers)) {
  for (d in 1:length(devices)) {
    
    diff_times %>%
      filter(scheduler == schedulers[[s]]) %>%
      filter(device_name == devices[[d]]) %>%
      select(device_id, device_name, plan_date_diff) -> one
    
    summary_row <- data.frame(matrix(ncol=length(col_names), nrow = 0))
    colnames(summary_row) <- col_names
    
    mean <- mean(one$plan_date_diff, na.rm = TRUE)
    summary_boxplot <- boxplot(one$plan_date_diff, plot = FALSE) 
    bottom_whisker <- summary_boxplot$stats[1,1]
    top_whisker <-  summary_boxplot$stats[5,1]
    
    summary_row[1,] <- c(schedulers[[s]], devices[[d]], fivenum(one$plan_date_diff), mean, bottom_whisker, top_whisker)

    summaries_diff <- rbind(summaries_diff, summary_row)
  }
}

summaries_diff <- 
  summaries_diff %>%
  mutate_at(vars(c("scheduler", "device_name")), ~as.character(.)) %>% 
  mutate_at(vars(-c("scheduler", "device_name")), ~as.numeric(.)) %>%
  mutate_if(is.numeric, ~round(., 3))

# glimpse(summaries_diff)

data %>%
  group_by(exp_id, scheduler, device_name) %>%
  summarise(n = n(),
            start_datetime = min(plan_date),
            end_datetime = max(plan_date)) %>%
  inner_join(summaries_diff, by=c("scheduler", "device_name")) %>%
  arrange(exp_id, scheduler, device_name) -> summaries_diff_table

kable(summaries_diff_table %>%
        select(`# exp` = exp_id,
               `Type scheduler` = scheduler,
               `Device` = device_name,
               `# records` = n,
               `Start` = start_datetime,
               `End` = end_datetime,
               `Min` = min,
               `Bottom whisker`= bottom_whisker,
               `1st Qu.` = quartile1,
               `Median` = median,
               `Mean` = mean,
               `3rd Qu.` = quartile3,
               `Top whisker`= top_whisker,
               `Max` = max),
      format = "html",
      booktabs = TRUE,
      caption = "Table. Descriptive statistics of 'planning timestamp(n-1) - planning timestamp(n)' [in order to detect wrong planning timestamp values] per scheduler/device.") %>%
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "bordered", "condensed")) %>%
  collapse_rows(columns = 1:2, valign = "top")
  # row_spec(6, color = "white", background = "orange") %>%
  # row_spec(9:10, color = "white", background = "#D7261E")

```




```{r plan_date_diff_outliers}
diff_times <-
  diff_times %>%
  left_join(summaries_diff, by=c("scheduler", "device_name")) %>%
  mutate(plan_date_diff_outlier_pos = ifelse(plan_date_diff > top_whisker, 1, 0)) %>%
  mutate(plan_date_diff_outlier_neg = ifelse(plan_date_diff < bottom_whisker, 1, 0)) %>%
  mutate(plan_date_diff_outlier = ifelse((plan_date_diff_outlier_pos == 1 | plan_date_diff_outlier_neg == 1), 1, 0)) %>%
  arrange(device_name, step)
```

```{r plan_date_diff_outliers_table}
plan_date_outliers <- 
  filter(diff_times, plan_date_diff_outlier==1) %>%
  group_by(scheduler, device_name) %>%
  summarise(n = n()) %>%
  arrange(device_name, scheduler)

plan_date_outliers_pos <- 
  filter(diff_times, plan_date_diff_outlier_pos==1) %>%
  group_by(scheduler, device_name) %>%
  summarise(n_pos = n()) %>%
  arrange(device_name, scheduler)

plan_date_outliers_neg <- 
  filter(diff_times, plan_date_diff_outlier_neg==1) %>%
  group_by(scheduler, device_name) %>%
  summarise(n_neg = n()) %>%
  arrange(device_name, scheduler)

plan_date_outliers <- 
  plan_date_outliers %>% 
  left_join(plan_date_outliers_pos, by = c("scheduler", "device_name")) %>%
  left_join(plan_date_outliers_neg, by = c("scheduler", "device_name"))
  

kable(plan_date_outliers %>%
        select(`Device` = device_name, 
               `Type scheduler` = scheduler,
               `# outliers` = n, 
               `# outliers positive` = n_pos,
               `# outliers negative` = n_neg),
      format = "html",
      booktabs = TRUE,
      caption = "Table. Outliers of consecutive planning timestamps per scheduler/device.") %>%
  kable_styling(full_width =FALSE,position = "left", bootstrap_options = c("striped", "bordered", "condensed"))
  # row_spec(c(8,10), color = "white", background = "#D7261E") %>%
  # row_spec(3, color = "white", background = "orange")

```




```{r plan_date_diff_outliers_plot, fig.height=16, fig.width=16}

diff_times %>%
  filter(plan_date_diff_outlier ==1, plan_date_diff < 200) %>%  #TODO: What do I do with super outliers?
  ggplot(aes(x = step, color=factor(device_name))) +
    # geom_point(aes(y = delay), alpha = 0.6, size = 0.5) +
    geom_point(aes(y = plan_date_diff), alpha = 0.3, size = 0.7) +
    # geom_point(aes(y = exec_date_diff)) +
  
    labs(title="plan_date[n-1] - plan_date[n]. without super-outliers (< 200 secs)",
         x = "Steps [~1min]") + 
    scale_y_continuous(name="diff [seconds]") + #, breaks=ybks_delay, limits=ylim_delay) +
    facet_grid(device_name~scheduler) +
    theme(legend.title = element_text(size=8), 
          legend.position="bottom",
          legend.direction = "horizontal") +
    theme_bw()

```




```{r plan_date_diff_zero_plot, fig.height=5, fig.width=12}

diff_times %>%
  filter(plan_date_diff == 0) %>%  
  ggplot() +
  geom_bar(aes(x=device_name, fill=scheduler), stat = 'count', position = position_dodge(preserve = 'single')) +
  theme_bw() +
  labs(title = "plan_date_diff == 0")

# 
# diff_times %>%
#   filter(plan_date_diff == 0) %>%
#   filter(device_id == names(devices)[4] | device_id == names(devices)[3]) %>%
#   ggplot(aes(x = plan_date, color=factor(device_name))) +
#     # geom_point(aes(y = delay), alpha = 0.6, size = 0.5) +
#     geom_point(aes(y = plan_date_diff), alpha = 0.3, size = 1) +
#     # geom_point(aes(y = exec_date_diff)) +
#   
#     labs(title="plan_date_diff == 0",
#          x = "Steps [~1min]") + 
#     scale_y_continuous(name="diff [seconds]") + #, breaks=ybks_delay, limits=ylim_delay) +
#     facet_grid(~device_name) +
#     theme(legend.title = element_text(size=8), 
#           legend.position="bottom",
#           legend.direction = "horizontal") +
#     theme_bw()


```







<!--
#####################################################
#####################################################
ABOVE THIS LINE IS NO LONGER VALID
#####################################################
#####################################################
-->







#### Execution timestamps (`delay2`)

- execution timestamp(n-1) - execution timestamp(n) gives us whether execution times betwewen consecutio tasks has been reliable. `delay2` substracts `r baseline_delay` to the previpus calculation.

```{r exec_dates_diff_summary}

diff_times <- 
  diff_times %>%
  group_by(scheduler, device_id) %>%
  arrange(exec_date) %>% # TODO: step
  mutate(exec_date_lagged = dplyr::lag(exec_date),
         exec_date_diff = as.numeric(as.duration(interval(exec_date_lagged, exec_date)) / dseconds(1)),
         exec_date_diff = ifelse(is.na(exec_date_diff), 0, round(exec_date_diff, 3)), 
         exec_date_diff_cum = cumsum(exec_date_diff))

# 'delay2' is the new metric
diff_times <- 
  diff_times %>%
  mutate(delay2 = exec_date_diff - baseline_delay)


col_names <- c("scheduler", "device_name", "min", "quartile1", "median", "quartile3", "max", "mean", "bottom_whisker", "top_whisker")
summaries_diff <- data.frame(matrix(ncol=length(col_names), nrow = 0))
colnames(summaries_diff) <- col_names

for (s in 1:length(schedulers)) {
  for (d in 1:length(devices)) {
    
    diff_times %>%
      filter(scheduler == schedulers[[s]]) %>%
      filter(device_name == devices[[d]]) %>%
      select(device_id, device_name, exec_date_diff) -> one
    
    summary_row <- data.frame(matrix(ncol=length(col_names), nrow = 0))
    colnames(summary_row) <- col_names
    
    mean <- mean(one$exec_date_diff, na.rm = TRUE)
    summary_boxplot <- boxplot(one$exec_date_diff, plot = FALSE) 
    bottom_whisker <- summary_boxplot$stats[1,1]
    top_whisker <-  summary_boxplot$stats[5,1]
    
    summary_row[1,] <- c(schedulers[[s]], devices[[d]], fivenum(one$exec_date_diff), mean, bottom_whisker, top_whisker)

    summaries_diff <- rbind(summaries_diff, summary_row)
  }
}

summaries_diff <- 
  summaries_diff %>%
  mutate_at(vars(c("scheduler", "device_name")), ~as.character(.)) %>% 
  mutate_at(vars(-c("scheduler", "device_name")), ~as.numeric(.)) %>%
  mutate_if(is.numeric, ~round(., 3))

# glimpse(summaries_diff)

data %>%
  group_by(exp_id, scheduler, device_name) %>%
  summarise(n = n(),
            start_datetime = min(plan_date),
            end_datetime = max(plan_date)) %>%
  inner_join(summaries_diff, by=c("scheduler", "device_name")) %>%
  arrange(exp_id, scheduler, device_name) -> summaries_diff_table

kable(summaries_diff_table %>%
        select(`# exp` = exp_id,
               `Type scheduler` = scheduler,
               `Device` = device_name,
               `# records` = n,
               `Start` = start_datetime,
               `End` = end_datetime,
               `Min` = min,
               `Bottom whisker`= bottom_whisker,
               `1st Qu.` = quartile1,
               `Median` = median,
               `Mean` = mean,
               `3rd Qu.` = quartile3,
               `Top whisker`= top_whisker,
               `Max` = max),
      format = "html",
      booktabs = TRUE,
      caption = "Table. Descriptive statistics of 'execution timestamp(n-1) - execution timestamp(n)' per scheduler/device.") %>%
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "bordered", "condensed")) %>%
  collapse_rows(columns = 1:2, valign = "top")
  # row_spec(6, color = "white", background = "orange") %>%
  # row_spec(9:10, color = "white", background = "#D7261E")

```


```{r fig.height=8, fig.width=12}

diff_times %>%
  filter(exec_date_diff < 200) %>%
  ggplot(aes(x = step, color=factor(device_name))) +
    # geom_point(aes(y = delay), alpha = 0.6, size = 0.5) +
    geom_point(aes(y = delay2), alpha = 0.3, size = 0.7) +

    labs(title="exec_date[n-1] - exec_date[n]",
         x = "Steps [~1min]") + 
    # scale_x_datetime(breaks = scales::date_breaks("12 hours"),
    #                  labels = scales::date_format("%d-%m %H:%M", tz="CET"),
    #                  expand = c(0,0)) +
    scale_y_continuous(name="diff [seconds]") + #, breaks=ybks_delay, limits=ylim_delay) +
    
    # guides(color=guide_legend(title="Scheduler",
    #                           # override.aes=list(fill=NA),
    #                           nrow=2)) + 
    # 
    # guides(color=FALSE) +
    facet_grid(device_name~scheduler) +
    theme(legend.title = element_text(size=8), 
          legend.position="bottom",
          legend.direction = "horizontal") +
    theme_bw()

```




```{r fig.height=8, fig.width=12}

diff_times %>%
  filter(device_id == names(devices)[5]) %>%
  ggplot(aes(x = step, color=factor(scheduler))) +
    # geom_point(aes(y = delay), alpha = 0.6, size = 0.5) +
    geom_point(aes(y = delay2), alpha = 0.3, size = 0.7) +

    labs(title= paste0(names(devices)[5], ": exec_date[n-1] - exec_date[n]"),
         x = "Steps [~1min]") + 
    # scale_x_datetime(breaks = scales::date_breaks("12 hours"),
    #                  labels = scales::date_format("%d-%m %H:%M", tz="CET"),
    #                  expand = c(0,0)) +
    scale_y_continuous(name="diff [seconds]") + #, breaks=ybks_delay, limits=ylim_delay) +
    
    # guides(color=guide_legend(title="Scheduler",
    #                           # override.aes=list(fill=NA),
    #                           nrow=2)) + 
    # 
    # guides(color=FALSE) +
    facet_wrap(~exec_day, nrow = 4) +
    theme(legend.title = element_text(size=8)) +
    theme_bw()

```



#### Framework overhead

- execution timestamp(n) - planning timestamp(n+1)


#### Delay

- (planning timestamp(n) - execution timestamp(n)) - 60




**BQ Aquaris V**

_How does `delay` behave in the middle 50% of data (IRQ)?_

Advance scheduler shows worse delay times. In addition, the central cluster shows steadily increased values of `delay` over time. In contrast, the basic scheduler shows uniform delays over time adn concentrated aoru 0.3 sec. Teh basic scheduler's boxplot shows a thinner box than the advance scheduler's. There seems no clear relationship between battery levels and delays, which suggests that low battery levels do not impact (or at least not significally) on the planned execution time of a task. 

```{r dual_scale_plots_BQ, fig.width=14, fig.height=10}
sel_device_name <- devices[[1]]
selection <- 
  data %>%
  filter(device_name == sel_device_name, outlier == 0) %>%
  arrange(plan_date)


plot_BQ_wo_outliers <- dual_scale_plot(selection, 
                           paste(sel_device_name, "(without outliers)"))

boxplot_BQ_wo_outliers <-
  filter(selection) %>%
    ggplot(aes(x = device_id, y = delay, fill=factor(exp_id))) + 
    geom_boxplot(alpha = 0.7) +
    facet_grid(~scheduler) +
    theme_bw() +
    # scale_fill_hue() +
    guides(fill=FALSE)

selection <- 
  data %>%
  filter(device_name == sel_device_name, outlier == 0, whisker == 0) %>%
  arrange(plan_date)


plot_BQ_wo_whisker <- dual_scale_plot(selection, 
                           paste(sel_device_name, "(without outliers & whiskers)"))

boxplot_BQ_wo_whisker <-
  filter(selection) %>%
    ggplot(aes(x = device_id, y = delay, fill=factor(exp_id))) + 
    geom_boxplot(alpha = 0.7,
                 outlier.colour = "#1F3552", outlier.shape = 20) +
    facet_grid(~scheduler) +
    theme_bw() +
    # scale_fill_hue() +
    guides(fill=FALSE)


ggarrange(plot_BQ_wo_outliers, boxplot_BQ_wo_outliers, 
          plot_BQ_wo_whisker, boxplot_BQ_wo_whisker, 
          widths = c(3, 1), heights = c(3, 3))

```


_What happens when battery is going to be fully recharged?_




**Honor 9**

_How does `delay` behave in the middle 50% of data (IRQ)?_


```{r dual_scale_plots_HO, fig.width=14, fig.height=8}

sel_device_name <- devices[[4]]
selection <- 
  data %>%
  filter(device_name == sel_device_name, outlier == 0) %>%
  arrange(plan_date)

plot_HO_wo_outliers <- dual_scale_plot(selection, 
                           paste(sel_device_name, "(without outliers)"))

selection <- 
  data %>%
  filter(device_name == sel_device_name, outlier == 0, whisker == 0) %>%
  arrange(plan_date)


plot_HO_wo_whisker <- dual_scale_plot(selection, 
                           paste(sel_device_name, "(without outliers & whiskers)"))


plot_HO_wo_outliers

plot_HO_wo_whisker
```

**Motorola Moto G**

```{r dual_scale_plots_MO, fig.width=14, fig.height=8}
sel_device_name <- devices[[5]]
selection <- 
  data %>%
  filter(device_name == sel_device_name, outlier == 0) %>%
  arrange(plan_date)

plot_MO_wo_outliers <- dual_scale_plot(selection, 
                           paste(sel_device_name, "(without outliers)"))

selection <- 
  data %>%
  filter(device_name == sel_device_name, outlier == 0, whisker == 0) %>%
  arrange(plan_date)


plot_MO_wo_whisker <- dual_scale_plot(selection, 
                           paste(sel_device_name, "(without outliers & whiskers)"))


plot_MO_wo_outliers

plot_MO_wo_whisker
```


See the [shiny app](https://cgranell.shinyapps.io/symptoms-scheduler/)! (TO BE UPDATED)






